\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2018-03-27 write_up/mixfishsim_ecolmod.tex
Change bib style of authoryear

%% Numbered
%\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
%%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
%\textcolor{gray}{[Guidance: A concise and factual abstract is required. The abstract 
%should state briefly the purpose of the research, the principal results and
%major conclusions. An abstract is often pnresented separately from the article,
%so it must be able to stand alone. For this reason, References should be
%avoided, but if essential, then cite the author(s) and year(s). Also,
%non-standard or uncommon abbreviations should be avoided, but if essential they
%must be defined at their first mention in the abstract itself.  Graphical
%abstract: Although a graphical abstract is optional, its use is encouraged as
%it draws more attention to the online article. The graphical abstract should
%summarize the contents of the article in a concise, pictorial form designed to
%capture the attention of a wide readership. Graphical abstracts should be
%submitted as a separate file in the online submission system. Image size:
%Please provide an image with a minimum of 531 X 1328 pixels (h X w) or
%proportionally more. The image should be readable at a size of 5 x 13 cm using
%a regular screen resolution of 96 dpi.  Preferred file types: TIFF, EPS, PDF or
%MS Office files.] \\}
%
Fishing exploits spatially and temporally hetergenous fish populations, using
species-unselective gear that can result in unintended, unwanted catch of low
quota or protected species. Reducing these unwanted catches is crucial for
biological and economic sustainability of `mixed fisheries' and implementation
of an ecosystem approach to fishing. \\

@@

Use of spatial management as a tool has been proposed as a method to reduce
discards. However, its implementation is hampered by lack of knowledge of fish
and fishery spatiotemporal dynamics and understanding of the scale at which
processes are important for management. Understanding the correct scale for
spatial management is crucial in order to implement measures at a resolution
that ensures effective management\cite{Dunn2016} while minimising economic
impact. For example, a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species. \\

Ensuring measures are implemented at an appropriate scale has been a challenge
in the past that has led to ineffectual measures with unitended consequences
such as limited impact towards the management objective or increased benthic
impact on previously unexploited areas (e.g. the cod closure in the North
Sea\cite{Rijnsdorp2001,Dinmore2003}). Since then more refined spatial
information has become available through the combination of logbook and Vessel
Monitoring System (VMS) data\cite{Lee2010, Bastardie2010, Gerritsen2012,
	Mateo2016} and more real-time spatial management has been possible 
	(e.g. \cite{Holmes2011}). Such information is, however, patchy and
derived from an inherently biased sampling programme (i.e. targeted fishing).
Further, fishers generally only recorded landings (not catch) on a daily basis.
This leads to questions about the validity of inference that can be drawn from
landings data assigned to VMS activity pings. \\

@@

%\textcolor{gray}{[Guidance: Provide sufficient details to allow the work to be
%reproduced by an independent researcher. Methods that are already published
%should be summarized, and indicated by a reference.  If quoting directly from a
%previously published method, use quotation marks and also cite the source. Any
%modifications to existing methods should also be described.] \\ }
%
We develop a simulation model with a modular event-based approach, where
modules are implemented on independent time-scales appropriate to capture the
characteristic of the process modelled (Figure 1). The fishing model operated on a
tow-by-tow basis, while population dynamics (fishing and natural mortality,
growth) operate on a daily time-step.  Population movement occurs on a
weekly time-step, while recruitment occurs periodcally each year for a set
time period (e.g. 3 weeks) at at specified point individual to a species. The
simulation framework is implemented in the statistical software package R
\cite{RCoreTeam2017}; available as an R package from the authors github
(\url{www.github.com/pdolder/MixFishSim}).\\

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \cite{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. Here, population biomass
growth and depletion for pre-recruits and fish recruited to the fishery are
modelled separately as a function of previous recruited biomass, intrinsic
population growth and recruitment:

@@

where $\rho$ is Brody's coefficient, shown to be approximately equal to
$exp(-K)$, where $K$ is the growth rate from a von bertalanffy logistic growth
model \cite{Schnute1985}. $Wt_{R-1}$ is the weight of fish prior to
recruitment, while $Wt_{R}$ is the recruited weight. $\alpha_{d}$ represents
the proportion of fish recruited during that day for the year, while
$R_{\tilde{y}}$ is the annual recruits. \\

Mortality $Z$ can be decomposed to natural mortality, $M$, and fishing
mortality, $F$, where both $M$ and $F$ are instantaneous rates with $M$ fixed
and $F$ calculated by solving the Baranov catch equation \cite{Hilborn1992b}
for $F$:

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{mixfishsim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
(\cite{Beverton1957}):

@@

or a stochastic Ricker form \cite{Ricker1954}

@@

For the habitat we define a Gaussian random field process, $\{S(x) : x \in
\mathbb{R}^2\}$, that is a stochastic process where any collection of locations
$x_{1}, \dots, x_{n}$ where for each $x_{i} \in \mathbb{R}^2$, the joint
distribution of $S = \{S(x1),\dots S(x_{n})\}$ is multivariate Gaussian. The
distribution is specified by its \textit{mean function}, $\mu(x) = E[S(x)]$ and
its \textit{covariance function}, $\gamma(x,x') = Cov\{S(x),S(x')\}$
\cite{Diggle2007}.\\

@@

In the simulation model, the habitat for each of the populations is generated
through the \textit{RFSimulate} function of the \textit{RandomFields} R package
\cite{Schlater2015}, implementing different parameter settings to affect the
patchiness of the populations. Each population is initialised at a single
location, and subsequently moves according to a probabilistic distribtion based
on habitat suitability, temperature and distance from current cell.

@@

Several studies (e.g.\cite{Hutton2004, Tidd2012, Girardin2015}) have confirmed
past activity and past catch rates are strong predictors of fishing location
choice. For this reason, the fleet dynamics sub-model includes a learning
component, where a vessel's initial fishing location in a trip is based on
selecting from previously successful fishing locations. This is achieved by
sorting all previous fishing events in the previous trip as well as the
previous time periods in past years, and choosing randomly from the top x \% of
fishing events in value.  Simulation testing indicated that this learning
increased the mean value of catches for the vessels, over just relying on the
correlated random walk function.

@@

Fishing locations within a trip are determined by a modified random walk
process. A random walk type was chosen as it is the simplest assumption
commonly used in ecology to describe animal movement which searching for
homogeneously distributed prey about which there is uncertain knowledge. In a
random walk, movement is a stochastic process through a series of steps that
can either be equal in length or take some other functional form.  The
direction of the random walk can be correlated, a characteristic known as
`persistence', providing some overall location of directional movement
\cite{Codling2008} or uncorrelated. \\

A \textit{lévy walk} is a particular form of random walk characterised by a
heavy-tailed distribution of step-length and has received a lot of attention in
ecological theory in recent years as having shown to have very similar
characteristics as those observed by animals in nature, and being a near
optimum searching strategy for predators pursuing patchily distributed prey
\cite{Bartumeus2005, Sims2008}.  \cite{Bertrand2007} showed that Peruvian
anchovy fishermen have a stochastic search pattern similar to that observed
with a lévy walk. However, it remains a subject of debate, with the contention
that search patterns may be more simply characteristed as random walks
\cite{Sakiyama2013} with specific patterns related to the characteristics of
the prey field \cite{Sims2012}. \\

@@

Where several fishing vessels are exploiting the same fish population
competition is known to play an important role in local distribution of fishing
effort \cite{Gillis1998}. If several vessels are fishing on the same patch of
fish, local depletion and interference will affect fishing location choice of
the fleet as a whole \cite{Rijnsdorp2000, Poos2007a}. In order to account for
this behaviour, the fishing sub-model operates spatially on a daily time-step
so that for future days the biomass available to the fishery is reduced in the
areas fished. The cumulative effect is to make heavily fished areas less
attractive as future fishing opportunities.

>>>>

%% Numbered
%\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
%%\bibliographystyle{elsarticle-num}
%%\bibliographystyle{elsarticle-harv}

%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
%\textcolor{gray}{[Guidance: A concise and factual abstract is required. The abstract 
%should state briefly the purpose of the research, the principal results and
%major conclusions. An abstract is often pnresented separately from the article,
%so it must be able to stand alone. For this reason, References should be
%avoided, but if essential, then citep the author(s) and year(s). Also,
%non-standard or uncommon abbreviations should be avoided, but if essential they
%must be defined at their first mention in the abstract itself.  Graphical
%abstract: Although a graphical abstract is optional, its use is encouraged as
%it draws more attention to the online article. The graphical abstract should
%summarize the contents of the article in a concise, pictorial form designed to
%capture the attention of a wide readership. Graphical abstracts should be
%submitted as a separate file in the online submission system. Image size:
%Please provide an image with a minimum of 531 X 1328 pixels (h X w) or
%proportionally more. The image should be readable at a size of 5 x 13 cm using
%a regular screen resolution of 96 dpi.  Preferred file types: TIFF, EPS, PDF or
%MS Office files.] \\}
%
Fishing exploits spatially and temporally hetergenous fish populations, using
species-unselective gear that can result in unintended, unwanted catch of low
quota or protected species. Reducing these unwanted catches is crucial for
biological and economic sustainability of `mixed fisheries' and implementation
of an ecosystem approach to fishing. \\

@@

Use of spatial management as a tool has been proposed as a method to reduce
discards. However, its implementation is hampered by lack of knowledge of fish
and fishery spatiotemporal dynamics and understanding of the scale at which
processes are important for management. Understanding the correct scale for
spatial management is crucial in order to implement measures at a resolution
that ensures effective management \citep{Dunn2016} while minimising economic
impact. For example, a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species. \\

Ensuring measures are implemented at an appropriate scale has been a challenge
in the past that has led to ineffectual measures with unitended consequences
such as limited impact towards the management objective or increased benthic
impact on previously unexploited areas (e.g. the cod closure in the North
Sea \citep{Rijnsdorp2001,Dinmore2003}). Since then more refined spatial
information has become available through the combination of logbook and Vessel
Monitoring System (VMS) data \citep{Lee2010, Bastardie2010, Gerritsen2012,
	Mateo2016} and more real-time spatial management has been possible 
	\citep[e.g.][]{Holmes2011}. Such information is, however, patchy and
derived from an inherently biased sampling programme (i.e. targeted fishing).
Further, fishers generally only recorded landings (not catch) on a daily basis.
This leads to questions about the validity of inference that can be drawn from
landings data assigned to VMS activity pings. \\

@@

%\textcolor{gray}{[Guidance: Provide sufficient details to allow the work to be
%reproduced by an independent researcher. Methods that are already published
%should be summarized, and indicated by a reference.  If quoting directly from a
%previously published method, use quotation marks and also citep the source. Any
%modifications to existing methods should also be described.] \\ }
%
We develop a simulation model with a modular event-based approach, where
modules are implemented on independent time-scales appropriate to capture the
characteristic of the process modelled (Figure 1). The fishing model operated on a
tow-by-tow basis, while population dynamics (fishing and natural mortality,
growth) operate on a daily time-step.  Population movement occurs on a
weekly time-step, while recruitment occurs periodcally each year for a set
time period (e.g. 3 weeks) at at specified point individual to a species. The
simulation framework is implemented in the statistical software package R
\citep{RCoreTeam2017}; available as an R package from the authors github
(\url{www.github.com/pdolder/MixFishSim}).\\

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. Here, population biomass
growth and depletion for pre-recruits and fish recruited to the fishery are
modelled separately as a function of previous recruited biomass, intrinsic
population growth and recruitment:

@@

where $\rho$ is Brody's coefficient, shown to be approximately equal to
$exp(-K)$, where $K$ is the growth rate from a von bertalanffy logistic growth
model \citep{Schnute1985}. $Wt_{R-1}$ is the weight of fish prior to
recruitment, while $Wt_{R}$ is the recruited weight. $\alpha_{d}$ represents
the proportion of fish recruited during that day for the year, while
$R_{\tilde{y}}$ is the annual recruits. \\

Mortality $Z$ can be decomposed to natural mortality, $M$, and fishing
mortality, $F$, where both $M$ and $F$ are instantaneous rates with $M$ fixed
and $F$ calculated by solving the Baranov catch equation \citep{Hilborn1992b}
for $F$:

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{mixfishsim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}:

@@

or a stochastic Ricker form \citep{Ricker1954}:

@@

For the habitat we define a Gaussian random field process, $\{S(x) : x \in
\mathbb{R}^2\}$, that is a stochastic process where any collection of locations
$x_{1}, \dots, x_{n}$ where for each $x_{i} \in \mathbb{R}^2$, the joint
distribution of $S = \{S(x1),\dots S(x_{n})\}$ is multivariate Gaussian. The
distribution is specified by its \textit{mean function}, $\mu(x) = E[S(x)]$ and
its \textit{covariance function}, $\gamma(x,x') = Cov\{S(x),S(x')\}$
\citep{Diggle2007}.\\

@@

In the simulation model, the habitat for each of the populations is generated
through the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, implementing different parameter settings to affect the
patchiness of the populations. Each population is initialised at a single
location, and subsequently moves according to a probabilistic distribtion based
on habitat suitability, temperature and distance from current cell.

@@

Several studies \citep[e.g.][]{Hutton2004, Tidd2012, Girardin2015} have confirmed
past activity and past catch rates are strong predictors of fishing location
choice. For this reason, the fleet dynamics sub-model includes a learning
component, where a vessel's initial fishing location in a trip is based on
selecting from previously successful fishing locations. This is achieved by
sorting all previous fishing events in the previous trip as well as the
previous time periods in past years, and choosing randomly from the top x \% of
fishing events in value.  Simulation testing indicated that this learning
increased the mean value of catches for the vessels, over just relying on the
correlated random walk function.

@@

Fishing locations within a trip are determined by a modified random walk
process. A random walk type was chosen as it is the simplest assumption
commonly used in ecology to describe animal movement which searching for
homogeneously distributed prey about which there is uncertain knowledge. In a
random walk, movement is a stochastic process through a series of steps that
can either be equal in length or take some other functional form.  The
direction of the random walk can be correlated, a characteristic known as
`persistence', providing some overall location of directional movement
\citep{Codling2008} or uncorrelated. \\

A \textit{lévy walk} is a particular form of random walk characterised by a
heavy-tailed distribution of step-length and has received a lot of attention in
ecological theory in recent years as having shown to have very similar
characteristics as those observed by animals in nature, and being a near
optimum searching strategy for predators pursuing patchily distributed prey
\citep{Bartumeus2005, Sims2008}.  \citet{Bertrand2007} showed that Peruvian
anchovy fishermen have a stochastic search pattern similar to that observed
with a lévy walk. However, it remains a subject of debate, with the contention
that search patterns may be more simply characteristed as random walks
\citep{Sakiyama2013} with specific patterns related to the characteristics of
the prey field \citep{Sims2012}. \\

@@

Where several fishing vessels are exploiting the same fish population
competition is known to play an important role in local distribution of fishing
effort \citep{Gillis1998}. If several vessels are fishing on the same patch of
fish, local depletion and interference will affect fishing location choice of
the fleet as a whole \citep{Rijnsdorp2000, Poos2007a}. In order to account for
this behaviour, the fishing sub-model operates spatially on a daily time-step
so that for future days the biomass available to the fishery is reduced in the
areas fished. The cumulative effect is to make heavily fished areas less
attractive as future fishing opportunities.
